---
title: "User's guide for `SeuratConverter`"
author:
- name: Kevin Rue-Albrecht
  affiliation: Kennedy Institute of Rheumatology, University of Oxford,
    Headington, Oxford OX3 7FY, UK.
  email: kevin.rue-albrecht@kennedy.ox.ac.uk
package: SeuratConverter
output:
  BiocStyle::html_document
abstract: |
  This package offers a simple conversion from the Seurat
  class to the SingleCellExperiment class.
vignette: |
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This package aims at making `seurat` objects (`r CRANpkg("Seurat")` package)
accessible for downstream
processing using common _Bioconductor_ methods and classes.
In particular, the most readily equivalent _Bioconductor_ class apt
to store the various components of a `seurat` object is the
`SingleCellExperiment` class (`r Biocpkg("SingleCellExperiment")` package).
Indeed, the `SingleCellExperiment` class contains slots to store:
* any number of _assays_, that encompasses the `@raw.data`, `@data`, and
    `@scaled.data` slots of a `seurat` object.
* sample meta-data, that corresponds to the `@meta.data` slot of a `seurat`
    object.
* feature meta-data, that encompasses the presence of individual genes in the
    `@var.genes` slot of a `seurat` object
* any number of reduced dimensions sets of coordinates, that are usually
    stored under the `@dr` slot of a `seurat` object

# Pre-requisite

## Installation

The `r Githubpkg("kevinrue/SeuratConverter")` package is currently hosted
on GitHub, and may be installed as follows:

```{r install, eval=FALSE}
devtools::install_github("kevinrue/SeuratConverter")
```

## Loading packages

Let us load the various packages required for this vignette;:

```{r require, message=FALSE}
stopifnot(
    require(Seurat),
    require(SingleCellExperiment),
    require(SeuratConverter)
)
```

# Examples

Let us load both `r CRANpkg("Seurat")` and `r Biocpkg("SingleCellExperiment")`
to set up the workspace

## Minimal `seurat` object

Let us use the demonstration object of the `r CRANpkg("Seurat")` package:

```{r pbmc_small}
pbmc_raw <- read.table(
  file = system.file('extdata', 'pbmc_raw.txt', package = 'Seurat'),
  as.is = TRUE
)
pbmc_small <- CreateSeuratObject(raw.data = pbmc_raw)
pbmc_small
```

This minimal `seurat` object can be converted to a `SingleCellExperiment`
as follows:

```{r pbmc_small_sce}
pbmc_sce <- as(pbmc_small, "SingleCellExperiment")
pbmc_sce
```

## Normalised `seurat` object

The `NormalizeData` function merely updates the `@data` slot of `seurat`
objects (while also storing the associated parameters in
the `@calc.params` slot):

```{r NormalizeData}
pbmc_small <- NormalizeData(object = pbmc_small)
names(pbmc_small@calc.params)
```

As a consequence, this `seurat` object can be converted to a
`SingleCellExperiment` without additional difficulty:

```{r NormalizeData_sce}
pbmc_sce <- as(pbmc_small, "SingleCellExperiment")
pbmc_sce
```

**Note:** the normalisation parameters stored in the `@calc.params` slot
are _not_ transferred to the `SingleCellExperiment` object.

## Scaled `seurat` object

The `ScaleData` function populates the `@scale.data` slot of `seurat`
objects (while also storing the associated parameters in
the `@calc.params` slot):

```{r ScaleData, message=FALSE}
pbmc_small <- ScaleData(
    object = pbmc_small, vars.to.regress = c("nUMI"), display.progress = FALSE)
dim(pbmc_small@scale.data)
```

The additional information in the `@scale.data` slot is stored as an additional
assay of the same name in the resulting `SingleCellExperiment` object:

```{r ScaleData_sce}
pbmc_sce <- as(pbmc_small, "SingleCellExperiment")
pbmc_sce
```

## Variable genes

The `FindVariableGenes` function populates the `@var.genes` and `@hvg.info`
slots of `seurat` objects (while also storing the associated parameters in
the `@calc.params` slot):

```{r FindVariableGenes}
pbmc_small <- FindVariableGenes(object = pbmc_small, do.plot = FALSE)
head(pbmc_small@var.genes)
```

The information in the `@hvg.info` slot, as well as the presence of individual
genes in the `@var.genes` slot are stored as columns
in the `@rowData` slot of the resulting `SingleCellExperiment` object:

```{r FindVariableGenes_sce}
pbmc_sce <- as(pbmc_small, "SingleCellExperiment")
pbmc_sce
```

## Coordinates in reduced dimensions

Functions such as `RunPCA` and `RunTSNE` add their respective coordinates
under the `@dr` slot (while also storing the associated parameters in
the `@calc.params` slot):

```{r RunPCA}
pbmc_small <- RunPCA(pbmc_small, do.print = FALSE)
names(pbmc_small@dr)
pbmc_small <-
    RunTSNE(pbmc_small, reduction.use = "pca", dims.use = 1:5, perplexity=10)
names(pbmc_small@dr)
```

Every cell embedding in the `@dr` slot will be transferred to the
`reducedDims` slot of the resulting `SingleCellExperiment` object:

```{r RunPCA_sce}
pbmc_sce <- as(pbmc_small, "SingleCellExperiment")
pbmc_sce
```

## CCA-aligned

The `RunCCA` function combines two `seurat` objects, and stores
the reduced dimensions obtained by canonical-correlation analysis (CCA)
under the `@dr` slot (while also storing the associated parameters in
the `@calc.params` slot).

```{r RunCCA}
pbmc1 <- SubsetData(pbmc_small,cells.use = pbmc_small@cell.names[1:40])
pbmc2 <- SubsetData(pbmc_small,cells.use = pbmc_small@cell.names[41:80])
pbmc1@meta.data$group <- "group1"
pbmc2@meta.data$group <- "group2"
pbmc_cca <- RunCCA(pbmc1, pbmc2)
pbmc_cca <- AlignSubspace(
    pbmc_cca,reduction.type = "cca", grouping.var = "group", dims.align = 1:2)
names(pbmc_cca@dr)
```

Naturally, PCA and t-SNE reduced dimensions from the original two `seurat`
objects are _not_ transferred to the new combined `seurat` objects.
If desired, they need to be re-computed on the new object:

```{r RunCCA_PCA}
pbmc_cca <- RunPCA(pbmc_cca, do.print = FALSE)
names(pbmc_cca@dr)
pbmc_cca <-
    RunTSNE(pbmc_cca, reduction.use = "pca", dims.use = 1:5, perplexity=10)
names(pbmc_cca@dr)
```

Overall, CCA-aligned reduced dimensions are treated no differently than
other dimensionality reduction embeddings; they are also transferred to
the `reducedDims` slot of the resulting `SingleCellExperiment` object:

```{r RunCCA_sce}
pbmc_sce <- as(pbmc_cca, "SingleCellExperiment")
pbmc_sce
```

# Session info {.unnumbered}

```{r sessionInfo, echo=FALSE}
sessionInfo()
```
